PubCrawl: A Publication Search Engine

### 2012-05-17 report/paper.tex
Combine design and implementation sections
Move overview of the system to the top, move PageRank stuff to that section

\section{PubCrawl Design}

\section{ACM Digital Library}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=5in]{images/component-graph}
  \end{center}
  \caption{Distribution of component sizes in the citation graph}
  \label{comp-size}
\end{figure}

@@

\subsection{Adapted PageRank}
\label{sec:pagerank}

PageRank is a well-known algorithm for estimating the importance of a web page
as a function of the number of incoming links \cite{brin98, page99}. It is
plausible to extend the algorithm by analogy to academic papers and citations.
However, applying PageRank directly to the ACM citation graph gives suboptimal
results due to the high number of small components. Papers in these components
end with disproportionately high PageRank because further iterations do not
change their initial value very much. Instead of a constant, we set the initial
value of a paper's PageRank to the proportion of all papers contained in its
component.

\section{Implementation}
\label{sec:implementation}
\subsection{Overview}
PubCrawl is an online search website built with the Django web framework backed by a Postgres database and Lucene search index. While the Postgres database stores metadata about papers such as the citation graph, author and venue information, Lucene stores indexes for paper titles, abstracts and fulltext. The PubCrawl interface allows users to submit search terms and relative weights to be assigned to title, abstract, full text searches and PageRank. Figure \ref{interface} shows a screenshot of the interface.

\begin{figure}[t]
\centering
	\includegraphics[width=8in]{images/interface}
	\caption{PubCrawl interface allowing search term and relative weight inputs.}	
	\label{interface}
\end{figure}

\subsection{Scores and Scaling}
PubCrawl combines four scores in all - a fulltext search score, a title search score, an abstract search score and the PageRank score. Since Lucene does't score on a pre-defined scale, we normalize all Lucene. Similarly, we normalize PageRanks. We normalize all scores on a per-query basis. This is particularly relevant to PageRank since results for a particular topic may all have low PageRank since the topic is studied by few researchers. By normalizing per-query, we do not penalize topics with a small number of papers.

\subsection{Ranking}
An important challenge faced by PubCrawl is how to effectively combine the four scores above to get an over quality score for a document. We implemented two types of simple ranking functions for this purpose:

\begin{itemize}
	\item 2-factor ranking: As a first step, we only allowed full text and page rank scores to be combined. We used a simple sorting technique for this since every document returned by the search has an associated full text and page rank score.
	\item 4-factor ranking: In order to combine multiple scores, i.e., to combine multiple sorted lists, we used the RankJoin algorithm. Briefly, RankJoin is a top-k algorithm that takes multiple sorted lists and traverses them in a round-robin fashion performing hash joins on already seen tuples. It is able to stop before traversing all tuples because once the max possible score of unseen tuples becomes lesser than the minimum score of the top-k tuples, we are guaranteed to have the required top-k set of tuples. We implemented a slight variation of RankJoin since we don't need joins across all 4 lists.
\end{itemize}

\subsection{Tuning Parameters}

To tune weights for the four factors, we obtained information about seminal papers on 6 topics from experts in the respective areas. We then manually picked weights that gave good results for all 6 topics.

\subsection{PageRank Parameters}
For PubCrawl's implementation of the adapted PageRank scoring function, we used several parameters for variables defined in Section \ref{sec:pagerank}:\\

\begin{itemize}
\item For the score source $E$, we use a uniform value. We minimize the value
  of $E$ that still results in convergence, to reduce the total amount of
  freely created score. In our system, we found $E=0.1$ to be appropriate.

\item For initializing node scores, we use a linear relationship with the size
  of the component containing the node. If the total number of papers (and
  nodes) is $N$ and paper $p$ is in a component with $X$ nodes, then the
  initial score for $p$ is set to $\frac{X}{N}$. Thus, larger components start
  with initially higher scores, and typically converge with higher final
  scores.

\item For the damping factor of our final scoring function (Equation
  \ref{eqn:damping}), we use $d=0.85$. This is a reflection of the PageRank
  study indicating this value as reasonable, rather than a precisely chosen
  one.

\end{itemize}

>>>>

\section{PubCrawl Design and Implementation}

\subsection{Overview}

PubCrawl is an online search website for the ACM Digital Library built with the
Django web framework. It is backed by a Postgres database which stores metadata
(authors, venues, and citations) and a Lucene search index. The PubCrawl
interface allows users to submit search terms and assign relative weights to
title, abstract, and full text searches and PageRank. Figure \ref{interface}
shows a screenshot of the interface.

\begin{figure}[bt]
  \begin{center}
    \includegraphics[width=8in]{images/interface}
  \end{center}
  \caption{PubCrawl user interface}
  \label{interface}
\end{figure}

\subsection{ACM Digital Library}

@@

\begin{figure}[tb]
  \begin{center}
    \includegraphics[width=5in]{images/component-graph}
  \end{center}
  \caption{Distribution of component sizes in the citation graph}
  \label{comp-size}
\end{figure}

@@

\subsection{Adapted PageRank}

PageRank is a well-known algorithm for estimating the importance of a web page
as a function of the number of incoming links \cite{brin98, page99}. It is
plausible to extend the algorithm by analogy to academic papers and citations.
We use the iterative algorithm as described but tune the parameters to improve
the quality of the results for the ACM data set.

We set the score source $E$ to a uniform vector with each coordinate equal to
$0.1$. This setting minimizes the PageRank each paper gets for free while still
allowing the algorithm to converge. We set the damping factor $d$ to the
suggested value of $0.85$. This setting reduces the number of iterations needed
to converge by reducing the amount of PageRank contributed in subsequent
iterations.

We found initializing the PageRank of each paper to a constant gave suboptimal
results because the ACM citation graph has a high number of small components.
Papers in these components end with disproportionately high PageRank because
further iterations do not change their initial value very much. Instead, we set
the initial value of a paper's PageRank to the proportion of all papers
contained in its component.

\subsection{Combining document scores and PageRanks}

We compute a final score used for ranking documents as a linear function of
subscores. One issue is that these subscores are not necessarily on the same
scale. In particular, Lucene scores are on an arbitrary scale. To address this
issue, we normalize scores per query.

The PageRank algorithm includes a normalization step at each iteration.
However, if a particular topic is studied by only a few researchers, all the
papers relevant to that topic will have low PageRank. To address this issue, we
renormalize PageRanks per query.

We implemented two ranking functions. The first only uses the full text score
and PageRank. The second uses a rank join--inspired algorithm to combine lists
of hits sorted by each of title, abstract, and full text scores and PageRank.
Briefly, the algorithm maintains the top $k$ hits seen so far. In each
iteration, we pop the maximum element off of the next list in round-robin order
and hash it. When we encounter a duplicate, we update the combined score and
its position in the top $k$. The algorithm terminates when the maximum possible
score of unseen tuples is less than the score of the $k$th hit.

To tune weights for the four factors, we obtained information about important
papers on 6 topics from experts in the respective areas. We then manually
picked weights that gave good results for all 6 topics.
